manuscript_figures

### 2024-05-29 scripts/manuscript_figures.Rmd
calculatted kelp biommass using morph and quad data

>>>>

```{r}
# read in quad data with filtered for just kelp
quad.filt <- QUAD %>%
        filter(SP_CODE == "SUGK")

# remove control column and two last columns, not needed
quad.filt <- quad.filt[ -c(13:14) ] 
quad.filt <- quad.filt[ -c(8) ] 
quad.filt <- quad.filt[ -c(5:6) ]

# I noticed that 06/30/23 dates were accidentily labeled 06/30/29
# Replace "29" with "23" in the DATE column
quad.filt$DATE <- sub("29$", "23", quad.filt$DATE)

# Convert DATE to Date format
quad.filt$DATE <- as.Date(quad.filt$DATE, format = "%m/%d/%y")

# Convert Date format to "YYYY-MM-DD"
quad.filt$DATE <- format(quad.filt$DATE, "%Y-%m-%d")

# combine quad data with morph data to calculate kelp biomass

#load in morph data and select for SUGK
morph.filt <- morph %>%
        filter(SP_CODE == "SL")

#rename SL to SUGK
morph.filt$SP_CODE[morph.filt$SP_CODE == "SL"] <- "SUGK"

# Convert the integer column to character
morph.filt$TRANSECT <- as.character(morph.filt$TRANSECT)

# merge morph filt and quad.filt to calculate biomass
merge.morph.quad <- left_join(morph.filt, quad.filt)

#notice some dates and data missing, Austin has to look this over!

#calculate kelp biomass
# To get wet weight in grams (biomass) you use the equation: y=0.00421*x^1.951 where x is kelp blade + stipe length.

# Calculate the kelp blade length
merge.morph.quad$kelp_blade_length <- merge.morph.quad$BLADE_WIDTH_CM * merge.morph.quad$BLADE_LENGTH_CM

# Calculate the sum of blade length and stipe length
merge.morph.quad$sum_length <- merge.morph.quad$kelp_blade_length + merge.morph.quad$STIPE_LENGTH_CM

# Define a function to calculate kelp biomass
calculate_kelp_biomass <- function(x) {
  y <- 0.00421 * x^1.951
  return(y)
}

# Calculate kelp biomass using the sum of blade length and stipe length
merge.morph.quad$kelp_biomass <- calculate_kelp_biomass(merge.morph.quad$sum_length)

# Remove the intermediate columns if no longer needed
merge.morph.quad$kelp_blade_length <- NULL
merge.morph.quad$sum_length <- NULL
